\section{Data sources}
In this project we had unprecedented collaborations with two satellite imagery providers, and a third automatic identifiaction system (AIS) data provider. High quality vessel data is essential to the success of a vessel risk framework, and hopefully this can be expanded in the future.

\subsection{Automatic identification system (AIS)}
Vessel location for the period from May 2016 to June 2017 was captured using the automatic identification system (AIS) transponders installed in every vessel with a length greater or equal to 30 meters. These transponders broadcast two types of messages. First, they transmit positional signals every $\approx$ 10 seconds, reporting GPS coordinates and navigational features of the vessel (i.e., speed, course, and heading). The second type of message is static and is reported every $\approx$ 30 seconds, these features include ship name, callsign, length, and port destination. 

Beacons near the shores capture these messages, working as an avoiding collision system with the sea coast and other vessels. This makes AIS an important component of maritime security \cite{Tetreault2005}. Coastal receivers do not always capture AIS signals away from shore.\footnote{According to Spire, the coastal receivers can only capture data in an 80 kilometers range.} Nonetheless, we use satellite captured AIS signal provided by Spire; this allows us to retrieve vessel tracking data from all over the world, and not only from vessels near seashores. Some caveats arise with the use of this source. First, AIS signals fetched by satellites can \textit{bunch together} leading to data loss, since the satellite is only able to process a limited number of signals at a time. Second, accuracy is not always certain. Hence there is a possible measurement error in the location and other features of the vessels. Lastly, there is the possibility of AIS tampering in vessels involved in illegal activities, such as IUU. 

\subsection{Satellite imagery}
To visually validate the AIS data, we used different sources of proprietary satellite imagery for the same timeframe of the positional AIS data. We rely on two sources of high-definition satellite imagery. The first source is Digital Globe (henceforth GBDX) which reports images with visual and Near Infrared (NIR) bands with a resolution of $\approx$ 3 meters per pixel. Second, we used imagery from Planet Labs, which has higher revisit times that GBDX, but at the cost of lower resolution (from $\approx$ 10 m. to $\approx$ 3 m. over the equator). 

Images from both sources were orthorectified and projected into a common grid. The first process corrects possible terrain distortions due to higher elevation angles \footnote{This angle is commonly known as Nadir. The terrain displacement occurs when satellite sensors are not orthogonal to the earth terrain. Thus the optimal Nadir angle is \ang{90}.} The second process projects the images into a common grid to georeference them into a common grid, usually under a \textit{UTM} local projection. Despite the availability of more advanced image processing tasks, as \textit{Pansharpening} or \textit{Atmospheric compensation}, we decided to use a single approach. Also, some advanced image processing tasks merge visual and NIR bands, hindering the use of a multispectral or thermal analysis. 


